Multiple imputation of incomplete multilevel data using Heckman selection models

### 2022-11-02 1.Paper/Paper.Rmd
Comments and suggestions for the entire paper. Due to a missing file I commented out the case study.

---
title: Multiple imputation of incomplete multilevel data using Heckman selection models
author:
- name: J. Munoz*
  num: a
- name: V. de Jong
  num: a
- name: T. Debray
  num: a
address:
- num: a
  org: Julius Center for Health Sciences and Primary Care, Utrecht Medical Center,Utrech, The Netherlands
corres: "* Johanna Munoz. \\email{j.munozavila@umcutrecht.nl}"
presentaddress: Julius Center for Health Sciences and Primary Care, Utrecht Medical Center, Universiteitsweg 100, 3584 CG Utrech, The Netherlands
authormark: Uthor \emph{et al}.
articletype: Research article
received: 2022-11-01
revised: 2022-12-01
accepted: 2023-01-01
abstract: "Missing data is a common problem in medical research, and is commonly addressed using multiple imputation. Although traditional imputation methods allow for valid statistical inference when data are missing at random (MAR), their implementation is not justified when observations are clustered (e.g., within studies) or when the presence of missingness depends on unobserved information. Unfortunately, this situation is increasingly common, and typically arises when individual participant data (IPD) are combined from multiple studies. While several imputation methods have been proposed for addressing individual studies where data are missing not at random (MNAR), their application and validity in large datasets with clustering remains unclear. We therefore explored the consequence of MNAR data in IPDMA in-depth, and proposed novel multilevel imputation methods for common missing patterns in clustered datasets. These methods build upon the principles of Heckman selection models, and adopt a two-stage meta-analysis approach for imputing binary and continuous variables. After evaluating the proposed imputation models in simulated scenarios, we illustrated their use in a malaria cross-sectional community survey to estimate the prevalence of parasitemia for children aged 2-10 years in five subregions in Uganda."
keywords: Heckman model; IPDMA; Missing not at random; Selection models;Multiple imputation;
bibliography: bibfile.bib
output: 
  rticles::sim_article:
  fig_caption: yes
  longtable: yes
  doublespace: yes

header-includes:
   - \usepackage{mathtools}
   - \usepackage{float}
   - \usepackage{multicol}
   - \usepackage{graphicx}
     \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
   - \usepackage{amsmath}
     \DeclareMathOperator\arctanh{arctanh}
---

# Introduction
Over the past few years, data sharing efforts have substantially increased and researchers increasingly often have access to large combined datasets derived from electronic health records or from individual participant data (IPD). For example, the clinical practice research datalink (CRPD) [@herrett2015] is an electronic health record (EHR) dataset in the UK, and has been used in a variety of medical research, such as the evaluation of health policy and drug efficacy. A recent example of an IPD-MA is the emerging risk factor collaboration [@theemergingriskfactorscollaboration2007], where data were combined from approximately 1.1 million individuals across 104 studies to investigate associations of cardiovascular diseases with several predictors. Individuals in these large datasets tend to be clustered in centers, countries or studies, where they have been subject to similar healthcare processes, and are therefore more alike than individuals from another cluster. Sometimes, clusters may also differ in participant eligibility criteria, follow-up length, predictor and outcome definitions, or in the quality of applied measurement methods. Hence, correlation is likely to be present between observations from the same cluster, which can lead to differences or ‘heterogeneity’ between clusters regarding baseline patient characteristics and subsequent outcomes.

Clustered datasets often contain  many incomplete variables. For example, in registry data it is common that test results are not available in registry data for all patients, as the decision to test may be at the discretion of the primary care physician or because the patient refuses to undergo testing . It is also possible that variables are systematically missing across clusters. For instance, in an IPD meta-analysis, it is common that studies collected information on different variables. Missing values may thus appear for all participants of a study in the combined dataset. The presence of missing data can lead to loss of statistical power, imbalance across clusters, bias in parameter estimates and therefore to erroneous conclusions as the analysis could be based on an unrepresentative sample.

To address the presence of missing data, it is important to consider  the proportion of missingness and the missing mechanism for each incomplete variable. Rubin (1976) [@rubin76] identified three missing mechanisms where the probability of missingness: 1) is constant (missing completely  at random; MCAR), 2) depends on observed data only (missing at random; MAR), or 3) depends on unobserved information even after conditioning on all observable variables (missing non-random; MNAR).  Traditional imputation methods are designed to address incomplete data sets where variables are MCAR or MAR. Their implementation is reasonable when there is no obvious mechanism of missingness, or when the observed data strongly relate to unobserved information.

Although it is possible to formally rule out MCAR [@little1988], it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms [@enders2022]. For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis. In practice, however, unless missingness is artificially introduced by study design, e.g. when a test is only taken on patients with certain characteristics, missingness will often (partially) depend on unobserved information.

One could see, a missing variable is a mixture of MAR and MNAR, depending on the available information, making the MAR mechanism more likely as more information is recorded. For instance, healthcare professionals may decide what type of measurement on indication, and depending on their personal experience or behavior decide whether or not to record the reasons for the measurement.

Registries are notoriously prone to incomplete variables that are MNAR, due to the complex recording process [@liu2021], e.g. laboratory tests are taken only in certain patients based on sign and symptom information that is often incomplete or not recorded. Also IPD may suffer from MNAR, for example when study participants who experience unfavorable results drop out of the study or also due to heterogeneity of the primary objective or resources of the studies involved, which may result in variables relevant to explain the missing process not being recorded.

The MNAR mechanism is considered as non-ignorable because to deal with this type of missingness it is necessary to specify information about the missingness process in addition to assumptions about the observed data. A common approach to address MNAR is to adopt  selection models such as the one proposed by Heckman (1976) [@heckman1976]. Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations, a main equation where the missing variable is associated with predictors, and a selection equation that accounts for the inclusion of observations in the sample.   An important feature of the Heckman selection model is that its implementation does not require data to be MNAR, and can also be used when data are MCAR or MAR. It therefore offers an appealing solution to incomplete data sets where the missingness mechanism is unknown.

Over the past few years, several Heckman selection models have been proposed for multiple imputation. Among them, Galimard et al.(2016) [@galimard_etal16] implemented a chained equations imputation method for continuous variables, which was extended to binary and categorical variables by employing copula estimates [@galimard_etal18].  Also, Ogundimu and Collins (2019) [@ogundimu_collins19] proposed a chained equations imputation method that is less dependent on Normality assumptions.

@@

Therefore, the aim of this paper is to develop a multilevel imputation method for incomplete continuous and binary variables that are MNAR. In section \ref{heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting. Then in the section \ref{methods} we define the main steps of the proposed imputation method. Posteriorly, in the section \ref{simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method. In section \ref{example} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children. 
Finally we provide a discussion  in section \ref{discussion} about  the results, limitations and propose future extension of our method.

@@

Selection bias occurs when the selection of a subject  or his observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause him or her to self-select out of the study or refuse to sign consent for a test), which in turn are correlated with unobserved variables related to a variable of interest (e.g., the result of a blood test) [@vella1998].

\begin{figure}[h!]
   \centering
   \includegraphics[scale=0.35]{DAG_heckman.png}
   \caption{DAG Heckman selection model: here the nodes (dotted = latent, continuous = observable) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent  selection variables}
   \label{fig:DAG}
\end{figure}
This can be better visualized in Figure \ref{fig:DAG}, where for the $j$th individual or unit within the $i$th cluster, there is $y_{ij}^*$, a latent outcome variable, and $r_{ij}^*$, a latent selection variable, which are correlated through $u_{ij}$ an unobserved or unrecorded variable, with $i\in[1,2,...,N]$ and $j\in[1,2,...,n_i]$. Here, both latent variables are related to $x_{ij}^O$ and $x_{ij}^S$ sets of predictor covariates. From $r_{ij}^*$ one can derive $r_{ij}=I(r_{ij}^*>=0)$ a selection indicator of $y_{ij}^*$ into the sample, and with this in turn, one can define $y_{ij}=y_{ij}^*,\forall r_{ij}=1$ the observable outcome variable.

Denoting $y^*_i=(y^*_{i1},y^*_{i2},...,y^*_{in_i})^T$ and  $r^*_i=(r^*_{i1},r^*_{i2},...,r^*_{in_i})^T$ the vectors of latent outcomes and latent selections in the $i$ cluster, then Heckman's model is defined by two main equations: the outcome equation (\ref{eq:1}), which describes the relation latent outcome-exposure, and the selection equation (\ref{eq:2}) which details the likelihood that the outcome is observed in the sample.
\begin{align}
y^*_{i}&=X^O_{i}\beta^O_{i}+\epsilon^O_{i}\label{eq:1}\\
r^*_{i}&=X^S_{ij}\beta^S_{i}+\epsilon^S_{i}\label{eq:2}
\end{align}

Here $\beta^O_{i}$ and $\beta^S_{i}$ are $p\times1$ and $q\times1$ coefficient parameter vectors and $\epsilon^O_{i}=(\epsilon^O_{i1},\epsilon^O_{i2},...,\epsilon^O_{in_i})^T$ and $\epsilon^S_{i}=(\epsilon^S_{i1},\epsilon^S_{i2},...,\epsilon^S_{in_i})^T$ are the residual terms vectors for the outcome and selection equations, respectively. Generally the same variables can be used on the matrix of predictor variables $X^O_{i}$ and $X^S_{i}$. However, to avoid multicollinearity problems [@puhani1997], it is recommended to include in $X^S_{i}$ at least one variable that is not included in the outcome model [@angrist2001]. This variable is commonly known as an exclusion restriction variable, and should only be associated with the selection in the sample $r^*_{ij}$ but not with the actual observation $y^*_{ij}$.

@@

Under the BVN assumption, the parameters of the Heckman model coefficients can be estimated using the two-step Heckman method (H2S) [@heckman1976] or the full information maximum likelihood method (FIML) [@amemiya1984]. However, both methods can lead to inconsistent estimators when this underlying distribution cannot be assumed[@gomes_etal20].To overcome this problem, other approaches have been proposed that relax the distribution assumptions, among them copula models [@smith03]. The copula approach uses a function, known as copula, that joins the marginal distribution of the error terms of the selection and outcome equation which are specified separately.
$$ F(r^*_{i},y^*_{i}) =\Phi\left(r^*_{i}-X^S_{i}\beta^S_{i},\frac{y^*_{i}-X^O_{i}\beta^O_{i}}{\sigma_i}, \rho_i\right)$$
Thus, to estimate the parameters of the Heckman method, it is sufficient to specify the marginal distributions of the error terms, and link them with a suitable copula function. In our imputation method we estimate the Heckman model using the copula method, as in real life many data could not follow a BVN distribution. This facilitates the Heckman model estimation and makes it more robust to deviations from the assumptions regarding the distribution of the data.

@@

Different procedures can be adopted to combine information between groups; however, in our imputation method we opted for the two-stage approach that is often used in meta-analyses. This is because such an approach is less computationally intensive and could potentially generate fewer convergence problems in the estimation of the Heckman hierarchical model compared to other approaches. Briefly, in a first stage $\theta_i$ the cluster specific parameters of the Heckman model,i.e., the parameters of the two equations in each study: the outcome model and the selection model, for each of the $N$ clusters are estimated separately and then, in a second stage, all $\theta_i$ are combined using a random effects model.

In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from an imaginary distribution of parameters with a $\theta_m$ population mean and a $\psi$ population variance [@higgins2009]. Thus, the  $\theta_i=\theta_m + b_i$ can be specified using $b_i \sim N(0,\psi)$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.

# Method
\label{methods}
We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data. Briefly, his method was developed to impute variables from a hierarchical structure (i.e., when there are unit samples grouped within a cluster or group).
This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random-effects meta-analysis.

With this method, values can be imputed in very common scenarios in IPD, e.g., sporadic and systematic missing patterns. In particular, when the response variable is systematically missing within a group, i.e., when $y_{ij}$ are totally missing within a group, the imputation values are drawn from a (generalized) linear model conditional on $\theta_m$ the marginal population parameters, i.e., those estimated after pooling the cluster-specific parameters $\theta_i$.On the other hand, when the variable is sporadically missing within the group, i.e., there are some observed $y_{ij}$ within the cluster, the imputation model is conditional on the shrunk-cluster parameters, i.e., those coming from the shrinkage of $\theta_i$ towards $\theta_m$.

Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster , thus obtaining $\theta_i=\{\beta_i^O,\beta_i^S,\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\theta_m$ parameter set at the marginal or population level. Our method is basically a univariate imputation method, but since it is implemented in a Gibbs sampling procedure, it can also be used to impute multiple incomplete variables in a data set.

@@

Initially, we estimate by the copula method the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with $y_{ij}^{obs}$ observable measurements within each $i$ group . 
The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.

@@

#### Fit a meta-analysis model
In this step, we pool the parameters $\widehat{\theta_i}$ with a random-effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the Heckman model.  In particular, we pooled the $p$ coefficients of the $\beta^O$ outcome equation and estimated a multivariate random effects meta-analysis model with them, similarly we combined all $q$ coefficient parameters of the $\beta^S$ selection equation. We also performed a univariate random effects meta_analysis on $\sigma'$ the log-transformed parameter of $\sigma$ and another on $\rho'$ the fisher-transformed parameter of $\rho$.

The meta-analysis model is performed with the \textbf{mixmeta} function of the R package mixmeta, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML) and moments estimation methods. For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between insensitivity and efficiency [@viechtbauer2005].

@@

#### Draw $\theta^*_i$ the cluster parameters

We draw $\theta^*_i$ shrunk-cluster parameters for each group $i$ from the following posterior distribution conditional on $\theta^*_m$ and $\psi^*$.
\begin{align*}
\theta^*_i\sim N\left(\frac{ \theta^*_m/\psi^*+\widehat{\theta_i}/\widehat{S_{\theta_i}}}{1/\psi^*+1/\widehat{S_{\theta_i}}},\frac{1}{1/\psi^*+1/\widehat{S_{\theta_i}}}\right)
\end{align*}

As can be seen, the mean and variance of the posterior distribution is a combination between the estimated marginal and cluster-specific parameters. Here the weights on the $\widehat{\theta_i}$ cluster-specific and the $\theta^*_m$ marginal parameters are inversely proportional to the $\widehat{S_{\theta_i}}$ within and $\psi^*$ between clusters variances. For example, when $\widehat{S_{\theta_i}} < \psi^*$ the mean of the conditional distribution gives more weight to the estimated cluster-specific parameter.
Conversely, when $\widehat{S_{\theta_i}}>\psi^*$, more weight is given to the estimated marginal parameters.  In the case of systematic missingness, it is like considering the within-cluster variance to be infinite ($\widehat{S_{\theta_i}}\to\infty$), then letting all parameters rely only on the marginal estimates.

>>>>

---
title: Multiple imputation of incomplete multilevel data using Heckman selection models
author:
- name: J. Munoz*
  num: a
- name: Valentijn M. T. de Jong
  num: a, b
- name: Thomas. P. A. Debray
  num: a
address:
- num: a
  org: Julius Center for Health Sciences and Primary Care, University Medical Center Utrecht, Utrecht University, Utrecht, The Netherlands
- num: b
  org: Data Analytics and Methods Task Force, European Medicines Agency, Amsterdam, the Netherlands
corres: "* Johanna Munoz. \\email{j.munozavila@umcutrecht.nl}"
presentaddress: Julius Center for Health Sciences and Primary Care, University Medical Center Utrecht, Universiteitsweg 100, 3584 CG Utrecht, the Netherlands
authormark: Uthor \emph{et al}.
articletype: Research article
received: 2022-11-01
revised: 2022-12-01
accepted: 2023-01-01
abstract: "Missing data is a common problem in medical research, and is commonly addressed using multiple imputation. Although traditional imputation methods allow for valid statistical inference when data are missing at random (MAR), their implementation is not justified when observations are clustered (e.g., within studies) or when the presence of missingness depends on unobserved information. Unfortunately, this situation is increasingly common, and typically arises when individual participant data (IPD) from multiple studies are combined. While several imputation methods have been proposed for addressing individual studies where data are missing not at random (MNAR), their application and validity in large datasets with clustering remains unclear. We therefore explored the consequence of MNAR data in IPD meta-analysis in-depth, and proposed novel multilevel imputation methods for common missing patterns in clustered datasets. These methods build upon the principles of Heckman selection models, and adopt a two-stage meta-analysis approach for imputing binary and continuous variables. After evaluating the proposed imputation models in simulated scenarios, we illustrated their use in a malaria cross-sectional community survey to estimate the prevalence of parasitemia for children aged 2-10 years in five subregions in Uganda."
keywords: Heckman model; IPDMA; Missing not at random; Selection models;Multiple imputation;
bibliography: bibfile.bib
output: 
  rticles::sim_article:
  fig_caption: yes
  longtable: yes
  doublespace: yes

header-includes:
   - \usepackage{mathtools}
   - \usepackage{float}
   - \usepackage{multicol}
   - \usepackage{graphicx}
     \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
   - \usepackage{amsmath}
     \DeclareMathOperator\arctanh{arctanh}
---
<!-- I tried to add a second affiliation for me, but I get 1,1 behind my name instead of 1,2. Do you happen to know how to fix this? -->
<!-- I find the last sentence of the abstract a bit hard to read. Does the following rephrasing make sense? -->
<!-- "..., we illustrated their use in a cross-sectional community survey to estimate the prevalence of malaria parasitemia in children aged 2-10 years in five subregions in Uganda."  -->

# Introduction
Over the past few years, data sharing efforts have substantially increased and researchers increasingly often have access to large combined datasets derived from electronic health records (EHR) or from individual participant data (IPD). For example, the clinical practice research datalink (CRPD) [@herrett2015] is an electronic health record dataset in the UK, which has been used in a variety of medical research, such as the evaluation of health policy and drug efficacy. A recent example of an IPD-MA is the emerging risk factor collaboration [@theemergingriskfactorscollaboration2007], where data were combined from approximately 1.1 million individuals across 104 studies to investigate associations of cardiovascular diseases with several predictors. Individuals in these large datasets tend to be clustered in centres, countries or studies, where they have been subject to similar healthcare processes, and are therefore more alike than individuals from another cluster. Sometimes, clusters may also differ in participant eligibility criteria, follow-up length, predictor and outcome definitions, or in the quality of applied measurement methods. Hence, correlation is likely to be present between observations from the same cluster, which can lead to differences or ‘heterogeneity’ between clusters regarding baseline patient characteristics and subsequent outcomes.

Clustered datasets often contain  many incomplete variables. For example, in registry data it is common that test results are not available in registry data for all patients, as the decision to test may be at the discretion of the primary care physician or because the patient refuses to undergo testing. It is also possible that variables are systematically missing across clusters. For instance, in an IPD meta-analysis, it is common that studies collected information on different variables. Missing values may thus appear for all participants of a study in the combined dataset. The presence of missing data can lead to loss of statistical power, imbalance across clusters, bias in parameter estimates and therefore to erroneous conclusions as the analysis could be based on an unrepresentative sample.

To address the presence of missing data, it is important to consider  the proportion of missingness and the missing mechanism for each incomplete variable. Rubin (1976) [@rubin76] identified three missing mechanisms where the probability of missingness: 1) is constant (missing completely  at random; MCAR), 2) depends on observed data only (missing at random; MAR), or 3) depends on unobserved information even after conditioning on all observable variables (missing not at random; MNAR).  Traditional imputation methods are designed to address incomplete data sets where variables are MCAR or MAR. Their implementation is reasonable when there is no obvious mechanism of missingness, or when the observed data strongly relate to unobserved information.

Although it is possible to formally rule out MCAR [@little1988], 
<!-- Stat med uses the AMA citation style. citations at the end of sentence go after the period: "out MCAR. [@little1988]" , and those at the end of a clause go after the period: "out MCAR, [@little1988]". Inline citations like "the clinical practice research datalink (CRPD) [@herrett2015]" are fine though.  See an example here:
https://onlinelibrary.wiley.com/doi/epdf/10.1002/sim.9221 -->
it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms [@enders2022]. For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis. In practice, however, unless missingness is artificially introduced by study design, e.g. when a test is only taken on patients with certain characteristics, missingness will often (partially) depend on unobserved information.

The missingness mechanism of a variable may be a mixture of MAR and MNAR, depending on the availability of data that may predict the missing data. Therefore, the MAR mechanism may become more influential than the MNAR mechanism as more information is recorded. <!-- MAR does not really become more likely. It is rather that the MAR mechanism becomes stronger than the MNAR mechanism. I have tried to capture this with the alternative text -->
For instance, healthcare professionals may decide what type of measurement on indication, and depending on their personal experience or behavior decide whether or not to record the reasons for the measurement.

Registries are notoriously prone to incomplete variables that are MNAR, due to the complex recording process. [@liu2021] e.g. laboratory tests are taken only in certain patients based on sign and symptom information that is often incomplete or not recorded. Also IPD may suffer from MNAR, for example when study participants who experience unfavorable results drop out of the study. Also, heterogeneity of the primary objective or resources of the studies involved may result in variables relevant to explain the missing process not being recorded.

The MNAR mechanism is considered as non-ignorable, because to deal with this type of missingness it is necessary to specify information about the missingness process in addition to assumptions about the observed data. A common approach to address MNAR is to adopt  selection models such as the one proposed by Heckman (1976) [@heckman1976]. Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations, a main equation where the missing variable is associated with predictors, and a selection equation that accounts for the inclusion of observations in the sample.   An important feature of the Heckman selection model is that its implementation does not require data to be MNAR, and can also be used when data are MCAR or MAR. It therefore offers an appealing solution to incomplete data sets where the missingness mechanism is not precisely known.

Over the past few years, several extensions and adaptations to the Heckman selection model have been proposed for multiple imputation. Among them, Galimard et al.(2016) [@galimard_etal16] implemented a chained equations imputation method for continuous variables, which was extended to binary and categorical variables by employing copula estimates [@galimard_etal18].  Also, Ogundimu and Collins (2019) [@ogundimu_collins19] proposed a chained equations imputation method that is less dependent on Normality assumptions.

@@

Therefore, the aim of this paper is to develop a multilevel imputation method for incomplete continuous and binary variables that are MNAR. In section \ref{heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting. Then in section \ref{methods} we define the main steps of the proposed imputation method. In section \ref{simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method. In section \ref{example} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children. 
Finally we provide a discussion  in section \ref{discussion} about  the results, limitations and propose future extension of our method.

@@

Selection bias occurs when the selection of a subject  or their observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause them to self-select out of the study or refuse to sign consent for a test), which in turn are correlated with unobserved variables related to a variable of interest (e.g., the result of a blood test) [@vella1998].

\begin{figure}[h!]
   \centering
   \includegraphics[scale=0.35]{DAG_heckman.png}
   \caption{DAG Heckman selection model: here the nodes (dotted = latent, continuous = observable) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent  selection variables}
   \label{fig:DAG}
\end{figure}
This is visualized in Figure \ref{fig:DAG}, where for the $j$th individual or unit within the $i$th cluster, there is $y_{ij}^*$, a latent outcome variable, and $r_{ij}^*$, a latent selection variable, which are correlated through $u_{ij}$ an unobserved or unrecorded variable, with $i\in[1,2,...,N]$ and $j\in[1,2,...,n_i]$. Here, both latent variables are related to the sets of predictor covariates $x_{ij}^O$ and $x_{ij}^S$. From $r_{ij}^*$ one can derive $r_{ij}=I(r_{ij}^*>=0)$ a selection indicator of $y_{ij}^*$ into the sample, and with this in turn, one can define $y_{ij}=y_{ij}^*,\forall r_{ij}=1$ the observable outcome variable.

Denoting $y^*_i=(y^*_{i1},y^*_{i2},...,y^*_{in_i})^T$ and  $r^*_i=(r^*_{i1},r^*_{i2},...,r^*_{in_i})^T$ the vectors of latent outcomes and latent selections in the cluster $i$, then Heckman's model is defined by two main equations: the outcome equation (\ref{eq:1}), which describes the relation latent outcome-exposure association, and the selection equation (\ref{eq:2}) which details the likelihood that the outcome is observed in the sample.
\begin{align}
y^*_{i}&=X^O_{i}\beta^O_{i}+\epsilon^O_{i}\label{eq:1}\\
r^*_{i}&=X^S_{ij}\beta^S_{i}+\epsilon^S_{i}\label{eq:2}
\end{align}

Here $\beta^O_{i}$ and $\beta^S_{i}$ are $p\times1$ and $q\times1$ coefficient parameter vectors and $\epsilon^O_{i}=(\epsilon^O_{i1},\epsilon^O_{i2},...,\epsilon^O_{in_i})^T$ and $\epsilon^S_{i}=(\epsilon^S_{i1},\epsilon^S_{i2},...,\epsilon^S_{in_i})^T$ are the residual terms vectors for the outcome and selection equations, respectively. Generally the same variables can be used on the matrix of predictor variables $X^O_{i}$ and $X^S_{i}$. However, to avoid multicollinearity problems [@puhani1997], it is recommended to include in $X^S_{i}$ at least one variable that is not included in the outcome model [@angrist2001]. This variable is commonly known as an exclusion restriction variable, and should only be associated with the selection in the sample $r^*_{ij}$ but not with the actual observation $y^*_{ij}$. 
<!-- Can you give some explanation (1-3 sentences?) on the exclusion and restriction assumptions here? -->

@@

Under the BVN assumption, the parameters of the Heckman model coefficients can be estimated using the two-step Heckman method (H2S) [@heckman1976] or the full information maximum likelihood method (FIML) [@amemiya1984]. However, both methods can lead to inconsistent estimators when this underlying distribution cannot be assumed[@gomes_etal20]. To overcome this problem, other approaches have been proposed that relax the distribution assumptions, among them copula models [@smith03]. The copula approach uses a function, known as a copula, that joins the marginal distribution of the error terms of the selection and outcome equation which are specified separately.
$$ F(r^*_{i},y^*_{i}) =\Phi\left(r^*_{i}-X^S_{i}\beta^S_{i},\frac{y^*_{i}-X^O_{i}\beta^O_{i}}{\sigma_i}, \rho_i\right)$$
Thus, to estimate the parameters of the Heckman method, it is sufficient to specify the marginal distributions of the error terms, and link them with a suitable copula function. In our imputation method we estimate the Heckman model using the copula method, as in real life many data could not follow a BVN distribution. This facilitates the Heckman model estimation and makes it more robust to deviations from the assumptions regarding the distribution of the data.

@@

Different procedures can be adopted to combine information between groups; however, in our imputation method we opted for the two-stage approach that is often used in meta-analyses. This is because such an approach is less computationally intensive and could potentially generate fewer convergence problems in the estimation of the Heckman hierarchical model compared to other approaches.

<!-- This sentence was too long and has the verbs all the way at the end, making it hard to read. I suggest to rephrase to three shorter sentences: -->
Briefly, in a first stage $\theta_i$ the cluster specific parameters of the Heckman model are estimated. That is, the parameters of the two equations in each study, the outcome model and the selection model, are estimated separately for each of the $N$ clusters. In the second stage, all $\theta_i$ are combined using a random effects meta-analysis model.

<!-- I wouldn't call the distribution imaginary. It is assumed to be real but latent. Some words we could use: conjectured, latent, presupposed, presumed  -->
In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from a latent distribution of parameters with a  population mean $\theta_m$ and a population variance $\psi$ [@higgins2009]. Thus, the  $\theta_i=\theta_m + b_i$ can be specified using $b_i \sim N(0,\psi)$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.

# Method
\label{methods}
We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data. Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are samples unit grouped within a cluster or group).
This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random effects meta-analysis.

With this method, values can be imputed in very common scenarios in IPD, e.g., sporadic and systematic missing patterns. In particular, when the response variable is systematically missing within a group, i.e., when $y_{ij}$ are totally missing within a group, the imputation values are drawn from a (generalized) linear model conditional on $\theta_m$ the marginal population parameters, i.e., those estimated after pooling the cluster-specific parameters $\theta_i$. On the other hand, when the variable is sporadically missing within the group, i.e., there are some observed $y_{ij}$ within the cluster, the imputation model is conditional on the shrunk-cluster parameters, i.e., those coming from the shrinkage of $\theta_i$ towards $\theta_m$.

Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster, thus obtaining $\theta_i=\{\beta_i^O,\beta_i^S,\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\theta_m$ parameter set at the marginal or population level. Our method is basically a univariate imputation method, but since it is implemented in a Gibbs sampling procedure, it can also be used to impute multiple incomplete variables in a data set.
<!-- I don't understand what you mean here: it is univariate and it can be used to impute multiple variables? I would say these two statements contradict each other. -->

@@

Initially, we use the copulat method to estimate the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each group $i$. 
The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.

<!-- Can you add a reference to the R package here? -->

@@

#### Fit a meta-analysis model
<!-- This paragraph starts of in the the present tense " we pool ", and later it's past tense " we pooled" etc. Stick to one tense. In applied papers, it is normal to use only past tense. In stats papers it is common to use present tense when describing a (new) method in the methods section. -->
In this step, we pool the parameters $\widehat{\theta_i}$ with a random effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the heckman model.  In particular, we pooled the $p$ coefficients of the $\beta^O$ outcome equation and estimated a multivariate random effects meta-analysis model with them, similarly we combined all $q$ coefficient parameters of the $\beta^S$ selection equation. 
<!-- The phrase " the log-transformed parameter of $\sigma$" is a description of $\sigma'$, so it is what is known as a commenting or parenthetical clause. That means you need to put a comma before and after this clause: "... on $\sigma'$, the log-transformed parameter of $\sigma$, ...". Although this article is not for the BMJ, I can recommend reading their house-style for writing: https://www.bmj.com/about-bmj/resources-authors/house-style -->
<!-- Read more here if you like: http://content.principia.edu/teaching-excellence/commenting-with-commas-when-and-where-to-use-them/ -->
We also performed a univariate random effects meta_analysis on $\sigma'$, the log-transformed parameter of $\sigma$, and another on $\rho'$, the fisher-transformed parameter of $\rho$.

The meta-analysis model is performed with the \textbf{mixmeta} function of the R package *mixmeta*, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods. For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between insensitivity and efficiency [@viechtbauer2005].
<!-- Insensitivity is not generally a good thing. I think you should either explain why it is a good thing here: "... it is insensitive to the true value of the mean of the distribution of random effects and it is efficient" (or something similar, feel free to rephrase) , or simplify it to "... a good balance between unbiasedness and efficiency."  -->

@@

#### Draw the cluster parameters $\theta^*_i$
<!-- Throughout the text, consider the order in which you put the symbol and the text: -->
<!-- Put the symbol after the text if the text is a description of the symbol: "shrunk-cluster parameters $\theta^*_i$" . -->
<!-- Put the text after the symbol if the symbol quantifies the text: "$n$ patients" , or " $N$ clusters" . -->
We draw the shrunk-cluster parameters $\theta^*_i$ for each group $i$ from the following posterior distribution conditional on $\theta^*_m$ and $\psi^*$.
\begin{align*}
\theta^*_i\sim N\left(\frac{ \theta^*_m/\psi^*+\widehat{\theta_i}/\widehat{S_{\theta_i}}}{1/\psi^*+1/\widehat{S_{\theta_i}}},\frac{1}{1/\psi^*+1/\widehat{S_{\theta_i}}}\right)
\end{align*}

As can be seen, the mean and variance of the posterior distribution is a combination between the estimated marginal and cluster-specific parameters. Here the weights on the cluster-specific parameters $\widehat{\theta_i}$ and the marginal parameters $\theta^*_m$ are inversely proportional to the within cluster variance $\widehat{S_{\theta_i}}$ and between clusters variance $\psi^*$. For example, when $\widehat{S_{\theta_i}} < \psi^*$ the mean of the conditional distribution gives more weight to the estimated cluster-specific parameter.
Conversely, when $\widehat{S_{\theta_i}}>\psi^*$, more weight is given to the estimated marginal parameters.  In the case of systematic missingness, it is like considering the within-cluster variance to be infinite ($\widehat{S_{\theta_i}}\to\infty$), then letting all parameters rely only on the marginal estimates. 
<!-- What does then mean in this last part? should it be " and lettting all parameters rely ..." ? -->
